Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}

### 2022-06-23 paper/paper1/ms.tex
Looking nice, about to print for some revisions

\documentclass[twocolumn]{aastex631}
\usepackage{hyperref}
\bibliographystyle{aasjournal}
%\turnoffedit
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{censor}
\usepackage{mathtools}
\usepackage{outlines}
\usepackage{amsmath,bm}
\definecolor{belize}{RGB}{41, 128, 185}
\definecolor{peter}{RGB}{52, 152, 219}
\definecolor{nephritis}{RGB}{39, 174, 96}
\definecolor{asbestos}{RGB}{127, 140, 141}
\hypersetup{linkcolor=asbestos, 
citecolor=belize,filecolor=asbestos,urlcolor=peter}

\hyphenation{long-it-udin-ally hy-po-thet-ical de-mon-stra-tion}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
    %\includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
    \includegraphics[width=0.45\textwidth]{figures/line_density_demo.png}
    \caption{Scaling of prominence and density of spectral lines.
        \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
        \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
    \label{fig_Nlines_vs_teff}
\end{figure}

@@

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo.png}
    \caption{PHOENIX spectrum cloned with blas\'e.  This $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity model has 9,028 individual cloned spectral lines, each with 3 tuned parameters.  The pictured 50 \AA~ chunk contains 121 spectral lines and represents about 1$\%$ of the entire spectral bandwidth that was cloned.  Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
    \label{fig_cloned_spectrum_demo}
\end{figure*}

@@

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/blase_clone_demo_zoom.png}
    \caption{Zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}. The cloned model has 7 spectral lines describing the 400 pixels in this 4 \AA~ chunk.  The native PHOENIX pixel sampling can be see as the boxy steps in both the native and cloned model.  }
    \label{fig_zoom_cloning_performance}
\end{figure*}

@@

\begin{eqnarray}
    \mathrm{regularization} \equiv -\ln{\mathrm{prior}}  = \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
\end{eqnarray}

@@

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_demo.png}
    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
    \label{fig_WASP69_demo}
\end{figure*}

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/blase_WASP69_regularized.png}
    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
    \label{fig_WASP69_transferred}
\end{figure*}

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/blase_super_resolution_template.png}
    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
    \label{fig_WASP69_regularized}
\end{figure*}

@@

\subsection{Prospects for closed-form computation of Jacobians}
In principle, one could compute the Jacobians analytically---since the partial derivatives of the Voigt function are known in closed-form.  This analytical approach would cut out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.  The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.  However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations that get applied at the joint telluric-and-stellar modeling step.  So analytic Jacobians could principally accelerate the cloning step but would add significant implementation complexity at the important data-model comparison step.  We, therefore, deem this extension somewhat diminishing returns on scarce developer time.

@@

\

\begin{acknowledgements}
    This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program, and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.  We acknowledge the National Science Foundation, which supported the work presented here under Grant No. 1910969.  This research has made use of NASA's Astrophysics Data System Bibliographic Services.
\end{acknowledgements}

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{hyperref}
\bibliographystyle{aasjournal}
%\turnoffedit
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{censor}
\usepackage{mathtools}
\usepackage{outlines}
\usepackage{amsmath,bm}

\definecolor{belize}{RGB}{41, 128, 185}
\definecolor{peter}{RGB}{52, 152, 219}
\definecolor{nephritis}{RGB}{39, 174, 96}
\definecolor{asbestos}{RGB}{127, 140, 141}
\hypersetup{linkcolor=asbestos, 
citecolor=belize,filecolor=asbestos,urlcolor=peter}

\hyphenation{long-it-udin-ally hy-po-thet-ical de-mon-stra-tion Ad-min-is-tra-tion}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
    %\includegraphics[width=0.45\textwidth]{figures/N_lines_vs_Teff_prom.png}
    \includegraphics[width=0.41\textwidth]{figures/line_density_demo.png}
    \caption{Scaling of prominence and density of spectral lines.
        \emph{Left:} Number of lines versus effective temperature for PHOENIX models truncated to IGRINS (\textbf{blue, connected} points) and HPF (\emph{red, free-standing} points) bandwidths, for different prominence thresholds of $0.02$, $0.01$, and $0.005$.
        \textbf{Right:} Number of lines per 100 \AA\ wavelength bin for stellar (\emph{blue, upper envelope of steps}) and telluric (\emph{orange, lower envelope of steps}), illustrated for a $T_\mathrm{eff}=4700$~K, $\log{g}=4.5$ PHOENIX model and a $T=290 \mathrm{K}$, relative humidity of 40\% TelFit model.}
    \label{fig_Nlines_vs_teff}
\end{figure}

@@

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/blase_clone_demo.png}
    \caption{PHOENIX spectrum cloned with blas\'e.  This $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity model has 9,028 individual cloned spectral lines, each with 3 tuned parameters.  The pictured 50 \AA~ chunk contains 121 spectral lines and represents about 1$\%$ of the entire spectral bandwidth that was cloned.  Some flaws can be seen near the cores of deep lines, or wings of broad lines.}
    \label{fig_cloned_spectrum_demo}
\end{figure*}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.48\textwidth]{figures/blase_clone_demo_zoom.png}
    \caption{Zoom in of the region between the vertical gray bars in Figure \ref{fig_cloned_spectrum_demo}. The cloned model has 7 spectral lines describing the 400 pixels in this 4 \AA~ chunk.  The native PHOENIX pixel sampling can be see as the boxy steps in both the native and cloned model.  }
    \label{fig_zoom_cloning_performance}
\end{figure}

@@

\begin{eqnarray}
    \mathrm{regularization} &\equiv& -\ln{\mathrm{prior}} \nonumber \\ 
    &=& \frac{1}{2} \sum_{j=1}^{N_\mathrm{lines}} \Big(\frac{\ln{a_j} - \ln{a^\prime_j }}{\Lambda}\Big)^2
\end{eqnarray}

@@

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_demo.png}
    \caption{WASP 69 observed with HPF, compared to a $T_{\mathrm{eff}}=4700\;$K,  $\log{g}=4.5$ solar metallicity PHOENIX model warped to $v\sin{i}=2.2$~km/s, $RV=-9.6$~km/s, and HPF resolving power.}
    \label{fig_WASP69_demo}
\end{figure*}

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/blase_WASP69_regularized.png}
    \caption{Semi-empirical model of WASP 69 transfer-learned with \texttt{blas\'e}, employing a regularization prior on the learned amplitudes.}
    \label{fig_WASP69_transferred}
\end{figure*}

\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/blase_super_resolution_template.png}
    \caption{The native-resolution semi-empirical model transfer-learned from WASP 69 HPF data. The revised model can be viewed as a super-resolution deconvolution of the HPF spectrum.}
    \label{fig_WASP69_regularized}
\end{figure*}

@@

\subsection{Prospects for closed-form computation of Jacobians}
In principle, one could compute the Jacobians analytically since the partial derivatives of the Voigt function are known in closed-form.  This analytical approach would cut out the autodiff approach, using the exact Jacobians to compute the Gradient Descent step.  The analytical approach may be faster and/or possess lower memory overhead than the autodiff one.  However, working out the chain-rule for the closed-form analytical Jacobians may become difficult or impossible once combined with the many augmentations that get applied at the joint telluric-and-stellar modeling step.  So analytic Jacobians could principally accelerate the cloning step but would add significant implementation complexity at the important data-model comparison step.  We, therefore, deem this extension somewhat diminishing returns on scarce developer time.

@@

\begin{acknowledgments}

These results are based on observations obtained with the Habitable-zone Planet Finder Spectrograph on the HET. The HPF team was supported by NSF grants AST-1006676, AST-1126413, AST-1310885, AST-1517592, AST-1310875, AST-1910954, AST-1907622, AST-1909506, ATI 2009889, ATI-2009982, and the NASA Astrobiology Institute (NNA09DA76A) in the pursuit of precision radial velocities in the NIR. The HPF team was also supported by the Heising-Simons Foundation via grant 2017-0494.

The Hobby-Eberly Telescope (HET) is a joint project of the University of Texas at Austin, the Pennsylvania State University, Ludwig-Maximilians-Universit\"at M\"unchen, and Georg-August-Universit\"at G\"ottingen. The HET is named in honor of its principal benefactors, William P. Hobby and Robert E. Eberly.

This material is based upon work supported by the National Aeronautics and Space Administration under Grant Numbers 80NSSC21K0650 for the NNH20ZDA001N-ADAP:D.2 program,
and 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.

We acknowledge the National Science Foundation, which supported the work presented here under Grant No. 1910969.

This research has made use of NASA's Astrophysics Data System Bibliographic Services.

\end{acknowledgments}
